---
title: ' Introduction to the Analysis of Sample Surveys with R - Exercise 1'
author: "Stefan Zins, Matthias Sand and Jan-Philipp Kolb"
date: "February 4th, 2016"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    keep_md: yes
---
*** 
1. Download the ESS dataset for [Sweden](http://www.europeansocialsurvey.org/data/country.html?c=sweden) (Survey Data and Sampling Design Data File (SDDF)) of the 5th round
2. Setup your workspace and load the R-packages [foreign](https://cran.r-project.org/web/packages/foreign/foreign.pdf) and [survey](https://cran.r-project.org/web/packages/survey/index.html)
3. Load the ESS dataset and the SDDF
4. Merge both data frames by their ID-variable, using the `merge()` command

***

5. Determine the sampling strategy (Inspect the variables `PSU`,`STRATFY` and `PROB`)
6. Add the variable `N` for the population size to your data frame. `N` can be calulated by
$$N= dweight* pweight *10000*n \text{,}$$
where $n$ refers to the sample size
7. Create a `svydesign` object from the dataset for Sweden using the `survey` package
8. Estimate the total and mean of the variable `tvtot`

*** 
# The survey package
- The survey package provides a large range of applications for complex survey samples
- Typically, the first step is to define a survey object with the `svydesign()` command

```{r echo=FALSE,message=FALSE}
library(Matrix)
library(survey)
```

### Simple Survey Object (Simple Random Sample)

```{r}
data(api)

surv.obj <- svydesign(id=~1,fpc = ~fpc, data = apisrs)
```

- `id` specifies the identifier of PSU and SSU;`id` $=$ \~ 0 or $=$ \~ 1 stipulates a single stage sampling
- For multi-stage samples the `id` argument should always specify a formula with the (cluster-) identifier at each stage
- `fpc` should be used for the finite population correction

      $\Rightarrow$  Either as the total population size of each stratum or as a fraction of the total population that has been sampled  
- `data` reflects the data set for which the design object should be defined

***

 
** ** | **Important Commands**
-----------|-----------------------------------------------------------------------------
`svytotal` |  returns the estimated total of a variable  and its standard error ($+ deff$)
`svymean` |  returns the estimated mean of a variable and its standard error ($+ deff$)
`svyquantile` |  Computes quantiles for data from complex surveys
`svyvar` |  Computes variances  for data from complex surveys
`weights` |  Returns the (design) weights of a survey object
`calibrate` | Calibration of a data set (uses the GREG-Estimator by default)



```{r}
svytotal(~api00,surv.obj)
```



