---
title: "Example 2"
author: "Jan-Philipp Kolb, Stefan Zins and Matthias Sand"
date: "1. Februar 2016"
output: md_document
---

```{r,echo=F}
Ex <- F
library(knitr)
```


## Download ESS

- Download the ESS dataset for [Denmark](http://www.europeansocialsurvey.org/file/download?f=ESS5DK.spss.zip&c=DK&y=2010) (Sampling Data and Country File) of the 5th round


## Packages for data import

- Use the package `foreign`

```{r,message=F,warning=F}
library(foreign)
```

```{r,message=F,warning=F}
library(memisc)
```

```{r,echo=F}
## Set working directory
main.path <- "J:/Work/Statistik/Kolb/Workshops/2016/grade_sampling/"

data.path <- paste(main.path,"data/",sep="")

setwd(data.path)
```


## [Import portable spss-files](http://stackoverflow.com/questions/3136293/read-spss-file-into-r)

```{r,eval=Ex}
DK <- as.data.set(spss.portable.file("ESS5DK.por"))
SE <- as.data.set(spss.portable.file("ESS5SE.por"))
```


```{r,eval=Ex}
DK <- as.data.frame(DK)
SE <- as.data.frame(SE)
```

```{r,eval=Ex,echo=F}

```

## The survey library

The dataset apistrat is a sample of schools from apipop
stratified by stype. 


Assuming the selection within the
strata was done by SRS, define a svydesign object that
enables your to make unbiased point and variance
estimates. Estimate the mean of api00.
3 Now you should try different allocations. Using stype
again as a stratification variable calculate the allocation of
a sample of 60 schools from apipop, using equal,
proportional to the number of schools, and optimal with
regard to api99 allocation.
4 Select a StrSRS from apipop for each of your allocations.
5 Estimate again the mean of api00 from your three different
samples.

## Load the ESS dataset and the country file
