---
title: "Sampling and Estimation - Exercise 2"
author: "Jan-Philipp Kolb, Stefan Zins and Matthias Sand"
date: "2 Februar 2016"
output: pdf_document
---


## Exercise 2.A

Estimation  under a stratified design

- [Download](https://github.com/BernStZi/SamplingAndEstimation/blob/short/tutorial/preparation/Preparation.md) the ESS for [Sweden](http://www.europeansocialsurvey.org/file/download?f=ESS7SE.spss.zip&c=SE&y=2010) and [Denmark](http://www.europeansocialsurvey.org/file/download?f=ESS5DK.spss.zip&c=DK&y=2010) (round 5)
- Import data to R and combine the two datasets
- Define a `survey` object (stratified design)
- Estimate the  empirical distribution of tv consumption (`tvtot`) in Sweden and Denmark 
- Estimate the joined empirical distribution of tv consumption in Sweden and Denmark


[Codebook ESS round 5](http://www.europeansocialsurvey.org/docs/round5/survey/ESS5_appendix_a6_e04_0.pdf)


## Exercise 2.B


- Load the `survey` package and the `api` datasets.

```{r,message=F}
library(survey)
data(api)
```


- The dataset `apistrat` is a sample of schools from `apipop` stratified by `stype`. Assuming the selection within the strata was done by SRS, define a survey object (`svydesign`) and calculate a point and variance estimate for the mean of `api00`. 

## Exercise 2.B

- Using `stype` again as a stratification variable try different allocations for stratified sample. Calculate the allocation of a sample of 60 schools from `apipop` using equal, proportional and optimal allocation. The proportional allocation should be proportional to the number of schools within the strata and the optimal alloaction should be optimal with regard to `api99`.

## Exercise 2.B

- Select a StrSRS from `apipop` for each allocation.
- Estimate again the mean of `api00` from all three samples and compare the results.



## Function for stratified samples

```{r}
strSRsample <- function(strind, nh, replace=FALSE){
  Nh <- table(strind)[names(nh)]
  h.id <- split(1:sum(Nh), strind)[names(nh)]
  
  
  sam <- mapply( function(x,y) sample(x, y, replace=replace)
                 , Nh, nh, SIMPLIFY = F)
  unlist(mapply(function(x,y) x[y]
                , h.id
                , sam, SIMPLIFY = F)
         ,use.names = FALSE)
}
```

## Getting the function

```{r,eval=F}
library(devtools)
install_github("BernStZi/SamplingAndEstimation/r/sampaest",
               ref="short")
```

```{r,eval=F}
url <- "https://raw.githubusercontent.com/BernStZi/
SamplingAndEstimation/short/r/sampaest/R/strSRsample.R"
source(url)
```

