---
title: "Sampling and Estimation - Exercise 2"
author: "Jan-Philipp Kolb, Stefan Zins and Matthias Sand"
date: "2 Februar 2016"
output: pdf_document
---


## Exercise 2.A

Estimation under a stratified design

- Download the ESS for Sweden and Denmark (round 5)
- Import data to R and combine the two datasets
- Define a `survey` object (stratified design)
- Estimate the  empirical distribution of tv consumption (`tvtot`) in Sweden and Denmark 
- Estimate the joined empirical distribution of tv consumption in Sweden and Denmark


## Solution Exercise 2.A 

- Download the ESS for [Sweden](http://www.europeansocialsurvey.org/file/download?f=ESS7SE.spss.zip&c=SE&y=2010) and [Denmark](http://www.europeansocialsurvey.org/file/download?f=ESS5DK.spss.zip&c=DK&y=2010) (round 5)


Use the package `foreign` for import

```{r,message=F,warning=F}
library(foreign)
```

```{r,echo=F}
Ex <- F
library(knitr)
set.seed(1133)
```

```{r,echo=F,eval=F}
main.path <- "C:/Users/kolbjp/Documents/GitHub/SamplingAndEsimation/tutorial"
setwd(main.path)
purl("Ex2solution.Rmd")
```


```{r,echo=F,eval=Ex}
## Set working directory
main.path <- "J:/Work/Statistik/Kolb/Workshops/2016/grade_sampling/"

data.path <- paste(main.path,"data/",sep="")

setwd(data.path)
```


Load the ESS dataset and the country file

```{r,eval=Ex}
DK <- read.spss("ESS5DK.sav",to.data.frame=T)
SE <- read.spss("ESS5SE.sav",to.data.frame=T)
```

## Construction of N

Later it is necessary for the specification of fpc.

```{r,eval=Ex}
DK$N <- DK$pweight*10000*nrow(DK)

SE$N <- SE$pweight*10000*nrow(SE)
```

## Combine the two datasets (DK,SE)

```{r,eval=Ex}
DK_tv <- data.frame(tvtot=as.character(DK$tvtot),
                    N=DK$N,
                    cntry=as.character(DK$cntry))
```

```{r,eval=Ex}
SE_tv <- data.frame(tvtot=as.character(SE$tvtot),
                    N=SE$N,
                    cntry=as.character(SE$cntry))
```

```{r,eval=Ex}
NE <- rbind(DK_tv,SE_tv)
```

## Adding new variables

```{r,eval=Ex}
NE$mt3 <- 0
NE$mt3[NE$tvtot=="More than 3 hours"] <- 1  
```



## R-package `survey`

```{r,message=F}
library(survey)
```

- [Define survey objects]((http://r-survey.r-forge.r-project.org/survey/example-design.html))

```{r,eval=Ex}
# SRS design
svydes_DK <- svydesign(id=~1,fpc=~N, data=DK)
svydes_SE <- svydesign(id=~1,fpc=~N, data=SE)
```

```{r,eval=Ex}
# Stratified design
svydes_NE <- svydesign(id=~1,strata=~cntry, 
                       fpc=~N, data=NE)
```



## The analysis functions in R-package `survey`

- Estimate the  empirical distribution of tv consumption (`tvtot`) in Sweden and Denmark 


```{r,eval=Ex}
stab_DK <- svytable(~tvtot,svydes_DK)
stab_DK
```


```{r,eval=Ex}
stab_SE <- svytable(~tvtot,svydes_SE)
stab_SE
```


## 

```{r,eval=Ex}
stab_NE <- svytable(~tvtot,svydes_NE)
stab_NE
```


## Visualisation

```{r,eval=Ex}
library(lattice)
barchart(stab_DK)
barchart(stab_SE)
```


## More analysis functions

- Estimate the joined empirical distribution of tv consumption in Sweden and Denmark

```{r,eval=Ex}
svytotal(~mt3,svydes_NE)

svymean(~mt3,svydes_NE)
```



## Exercise 2.B


- Load the `survey` package and the `api` datasets.

```{r,message=F}
library(survey)
data(api)
```

## The dataset [apipop](http://www.inside-r.org/packages/cran/survey/docs/api)

- The dataset `apistrat` is a sample of schools from `apipop` stratified by `stype`. 

```{r,echo=F}
kable(apipop[1:8,1:5])
```




## Computing the mean with `survey` package

- Assuming the selection within the strata was done by SRS, define a survey object (`svydesign`) and calculate a point and variance estimate for the mean of `api00`. 

```{r}
svy_api <- svydesign(id=~1,strata=~stype, 
                       fpc=~fpc, data=apistrat)

svymean(~api00,svy_api)
```


## Exercise 2.B

- Using `stype` again as a stratification variable try different allocations for stratified sample. Calculate the allocation of a sample of 60 schools from `apipop` using equal, proportional and optimal allocation. The proportional allocation should be proportional to the number of schools within the strata and the optimal alloaction should be optimal with regard to `api99`.

## Function for stratified samples

```{r}
strSRsample <- function(strind, nh, replace=FALSE){
  Nh <- table(strind)[names(nh)]
  h.id <- split(1:sum(Nh), strind)[names(nh)]
  
  
  sam <- mapply( function(x,y) sample(x, y, replace=replace)
                 , Nh, nh, SIMPLIFY = F)
  unlist(mapply(function(x,y) x[y]
                , h.id
                , sam, SIMPLIFY = F)
         ,use.names = FALSE)
}
```

## Getting the function

```{r,eval=F}
library(devtools)
install_github("BernStZi/SamplingAndEstimation/r/sampaest",
               ref="short")
```

```{r,eval=F}
url <- "http://raw.githubusercontent.com/BernStZi/
SamplingAndEstimation/short/r/sampaest/R/strSRsample.R"
source(url)
```


## Equal allocation:

```{r}
nh.eq <- c(20,20,20)
names(nh.eq) <- names(table(apipop$stype))

s.eq <- strSRsample(apipop$stype, nh.eq, replace=FALSE)
```

## Proportional allocation:

```{r}
Nh.tab  <- table(apipop$stype)
n <- 60
nh.pr <- round(Nh.tab/sum(Nh.tab)*n)

s.pr <- strSRsample(apipop$stype, nh.pr, replace=FALSE)
```

## Optimal allocation: 

```{r,echo=F,eval=T}
V.h   <- tapply(apipop$api99,apipop$stype,sd)[names(Nh.tab)]
nh.op <- round((Nh.tab*V.h)/(sum(Nh.tab*V.h))*n)
```


```{r}
s.op <- strSRsample(apipop$stype, nh.op, replace=FALSE)
```


## Exercise 2.B

- Select a StrSRS from `apipop` for each allocation.


Subselect the sample:

```{r}
pop <- apipop
pop$Nh <- table(apipop$stype)[apipop$stype]
strSRS.eq <- pop[s.eq,]
strSRS.pr <- pop[s.pr,]
strSRS.op <- pop[s.op,]
```

## `survey` - equal allocation

- Estimate again the mean of `api00` from all three samples and compare the results.

```{r}
svystrSRS.eq <- svydesign(ids=~cds, strata =~stype,  
                          fpc=~Nh, data=strSRS.eq)

svymean(~api00, svystrSRS.eq)
```

## `survey` - proportional allocation

```{r}
svystrSRS.pr <- svydesign(ids=~cds, strata =~stype,  
                          fpc=~Nh, data=strSRS.pr)

svymean(~api00, svystrSRS.pr)
```


## `survey` - optimal allocation

```{r}
svystrSRS.op <- svydesign(ids=~cds, strata =~stype,  
                          fpc=~Nh, data=strSRS.op)

svymean(~api00, svystrSRS.op)
```




