---
title: "Sampling and Estimation - Exercise 2"
author: "Jan-Philipp Kolb, Stefan Zins and Matthias Sand"
date: "2 Februar 2016"
output: beamer_presentation
---


## Exercise 2.A

Estimation under a stratified design

- Download the ESS for Sweden and Denmark (round 5)
- Import data to R and combine the two datasets
- Define a `survey` object (stratified design)
- Calculate the combined total for the tv consumption (`tvtot`) and compare it with the totals in Sweden and Denmark

## Solution Exercise 2.A 

- Download the ESS for Sweden and Denmark (round 5)

Download the ESS dataset for [Denmark](http://www.europeansocialsurvey.org/file/download?f=ESS5DK.spss.zip&c=DK&y=2010) (Sampling Data and Country File) and [Sweden](http://www.europeansocialsurvey.org/file/download?f=ESS7SE.spss.zip&c=SE&y=2010) of the 5th round

Use the package `foreign` for import

```{r,message=F,warning=F}
library(foreign)
```


```{r,echo=F,eval=Ex}
## Set working directory
main.path <- "J:/Work/Statistik/Kolb/Workshops/2016/grade_sampling/"

data.path <- paste(main.path,"data/",sep="")

setwd(data.path)
```


Load the ESS dataset and the country file

```{r,eval=Ex}
library(foreign)
DK <- read.spss("ESS5DK.sav",to.data.frame=T)
SE <- read.spss("ESS5SE.sav",to.data.frame=T)
```


## Exercise 2.B


- Load the `survey` package and the `api` datasets.

```{r,message=F}
library(survey)
data(api)
```


- The dataset `apistrat` is a sample of schools from `apipop` stratified by `stype`. Assuming the selection within the strata was done by SRS, define a survey object (`svydesign`) and calculate a point and variance estimate for the mean of `api00`. 

## Exercise 2.B

- Using `stype` again as a stratification variable try different allocations for stratified sample. Calculate the allocation of a sample of 60 schools from `apipop` using equal, proportional and optimal allocation. The proportional allocation should be proportional to the number of schools within the strata and the optimal alloaction should be optimal with regard to `api99`.

## Exercise 2.B

- Select a StrSRS from `apipop` for each allocation.
- Estimate again the mean of `api00` from all three samples and compare the results.



